\chapter*{Acknowledgements}
\markboth{Acknowledgements}{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
% put your text here
The work completed for this project would not have been possible without the assistance, help and support from several individuals. I would first like to thank Prof. Joseph Labuz for agreeing to be my thesis supervisor and giving me the chance to conduct this project at the University of Minnesota. He welcomed me with open arms and his door was always open whenever I had a question. His precious guidance and insightful comments helped me in all the time of research and writing of this thesis.  \newline

I would also like to express my sincere gratitude to Prof. Lyesse Laloui for his help and unfailing support all along my master studies at EPFL, and particularly in the search of a master thesis abroad. He always took the time to guide and help me in doubt times on my future perspectives.\newline

I thank Pouyan Asem for his contribution to this thesis, without whom the experiments performed in this work would not have been possible. I really enjoyed our stimulating discussions which created a great atmosphere in the rock mechanics lab, and his review of this thesis was greatly appreciated. I also thank Chen, Xiaowan and Mugurel whose help was important to the success of the experiments.\newline

My stay in the beautiful city of Minneapolis would not have been the same without Vicki. She welcomed me like I was part of her family and took care of me during these five month away from everything I knew, and I would be forever grateful for having the chance to meet her.\newline

Last but no least, I would like to thank my family, my brother and my parents for their unconditional support and  constant encouragement. Thank you to Robin for his support, encouragement, patience, catching enthusiasm and for teaching me most of the programming skills used in this thesis. 

\bigskip
 
\noindent\textit{Minneapolis, January $16^{th}$ 2020}
\hfill H.~F.
